\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{url}

\begin{document}

\begin{center}
\textbf {Readings for ENGL 516X, Spring Semester 2018}
\end{center}

Here are some of the resources to help you learn Python. Note: I may update this as time comes. 

\begin{itemize}
\item Primary textbook: Python for Informatics by Charles Severence
   \begin{itemize}
   \item can be purchased on amazon.com. Free ebook is available at: \\ \url{https://py4e.com/book.php}
   \item py4e.com website has several other useful materials.
   \end{itemize}
 \item Secondary textbook: How to think like a computer scientist: Learning with Python 3
 \begin{itemize}
 \item Freely available at: \url{http://openbookproject.net/thinkcs/python/english3e/} and several other free online versions exist.
 \item Interactive version: \url{http://interactivepython.org/runestone/static/thinkcspy/index.html}
 \item Also available for purchase on amazon.com
 \end{itemize} 
 \item For restless and hyperactive folks who are ahead of others in learning Python: Natural Language Processing with Python: \url{http://www.nltk.org/book/} - I will deal with the contents of this book in 520 in Fall 2018.

\item Useful websites:
   \begin{itemize}
     \item python.org - everything you want related to Python
   \item https://www.learnpython.org/ - interactive tutorials.
   \item  pythontutor.com  - website to visualize what happens when your code "runs"
      \end{itemize}
     
     \item online courses in Coursera by Charles Severence, under "Python for Everybody" specialization series - these used to be free in 2016. You can perhaps still view lectures for free. 
   \begin{itemize}
    \item Course 1: Getting Started with Python. \\ \url{https://www.coursera.org/learn/python}
    \item Course 2: Python Data Structures. \\ \url{https://www.coursera.org/learn/python-data}
    \item Course 2: Using Python to access Web Data. \\ \url{https://www.coursera.org/learn/python-network-data}
    \item Course 4: Using Databases with Python. \\ \url{https://www.coursera.org/learn/python-databases}
   \end{itemize}
   
   \item Other useful readings:
   \begin{itemize}
    \item "A Byte of Python" by C H Swaroop. \\ \url{http://www.swaroopch.com/notes/python/#frontpage}
    \item Python for Linguists by Ron Zacharski. \\ \url{http://www.zacharski.org/python2/} (This is very old - syntax changes happened!)
   \item Unix for Poets by Kenneth Ward Church. \\ \url{http://web.stanford.edu/class/cs124/kwc-unix-for-poets.pdf}
   \end{itemize}
  
  \item Other books:
  \begin{itemize}
  \item Dive into Python: \url{http://www.diveintopython3.net/}
   \item Learn Python the Hard Way (not free): \url{https://learnpythonthehardway.org/}
   \item For a more general introduction to computing: "Introduction to Computing: Explorations in Language, Logic and Machines" by David Evans. Can be purchased or downloaded at: \\ \url{http://www.computingbook.org/} (Difficult!)
  \end{itemize}


\end{itemize}
\end{document}
