\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage{savesym}
\savesymbol{checkmark}
\usepackage{dingbat}

\author[Sowmya Vajjala]{Instructor: Sowmya Vajjala}


\title[ENGL 516X]{ENGL 516X: \\ Methods of Formal Linguistic Analysis}
\subtitle{Semester: Spring '18}

\date{8 March 2018}

\institute{Iowa State University, USA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frame}\titlepage
\end{frame}

\begin{frame}%2minutes
\frametitle{Class outline}
\begin{itemize}
\item Assignment 5 description
\item Tuesday's last question
\item Final project ideas and discussion
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Notes from Midterm feedback-1}
Delta for me
\begin{itemize}
\item More time to practice step by step (\leftthumbsup)
\item More explanations needed, more repetitions needed (\leftthumbsup)
\item Clearer instructions about Assignments (\leftthumbsup)
\item More real life examples (\leftthumbsup)
\item Assignments don't coincide with what is taught in class (Really?)
\item Group work from beginning
\item Exercise time spread out in the class, not in the end.
\item Let us choose groups 
\item Flipped classroom
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Notes from Midterm feedback-2}
Delta for you
\begin{itemize}
\item Practice more outside class
\item Work on the program a little bit before searching internet for help
\item Office hours
\item Participate more in the forum
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Discussion on Tuesday's last question}
Write a small program now using regex, which takes a word as input and returns the plural word as output. Here are the rules to code:
\begin{itemize}
\item If a word ends in s, x, or z, add es to the end of the word.
\item If a word ends in a consonant +y, add ies to the word (removing y). If a word ends in a vowel+y, add s in the end (vacancy is vacancies but day is days) 
\item If none of the above cases are valid for a word, just add s at the end of the word and be happy with that. 
\end{itemize}
(Your input should not be a number or a string with numbers, punctuation etc. It has to be an alphabetic string.)
\end{frame}

\begin{frame}
\frametitle{}
Your solutions - Discussion
\end{frame}

\begin{frame}[fragile]
\frametitle{Taichi's solution}
\footnotesize
\begin{verbatim}
import re

word = input('Enter a word in its singular form: ')
plural = ""

if re.match("\w+(s|x|z)",word):
    plural = word + "es"
elif re.match("\w+(a|e|i|o|u)y", word):
    plural = word + "s"
elif re.match("\w+y", word):
    word_list = list(word)
    word_list[-1] = "i"
    word = "".join(word_list)
    plural = word + "es"
else:
    plural = word + "s"

print(plural)
\end{verbatim}
- What do you think this prints, if I give "boxy" as input? (expected output: boxies)
\end{frame}

\begin{frame}[fragile]
\frametitle{Emily's solution}
\scriptsize
\begin{verbatim}
import re
try:
    user_input = str(input("Please enter a word to make it plural: "))
    if user_input.isalpha():
        if re.search("[sxz]$", user_input):
            plural = user_input + "es"
            print(plural)
        elif re.search("[qwrtpsdfghjklzxcvbnm]y$", user_input):
            user_input.split()
            plural = user_input[0:-1] + "ies"
            print(plural)
        else:
            print(user_input + "s")
    else:
        print("Please enter a string.")
except:
    print("There was an error.")
\end{verbatim}
(My addition: in your elif, you can put "[\^{}aeiou]" instead. I don't understand the split() call though. 
\end{frame}

\begin{frame}[fragile]
\frametitle{Fatemeh's solution}
\tiny
\begin{verbatim}
import re
while True:
    try:
            word= str(input("please enter a word: "))
            if  word.isalpha():
                if re.search("\w+(s|x|z)", word):
                 word = word+"es"
                elif re.search("\w+(b|c|d|f|g|h|j|k|l|m|n|p|q|r|t|s|v|w|x|y|z)y", word):
                    word = word[0:-1] + "ies"
                elif re.search("\w+(a|e|o|u|i)y", word):
                    word = word + "s"
                else:
                     word = word +"s"
                print (word)
            else:
                print("the input is not valid")
    except:
        print( "please enter a word: ")
        continue
\end{verbatim} \small
Comments: Same issue as Taichi's. In Regular expression, you can use square brackets, and \^{} symbol for elegance in code. But, does this code ever stop?
\end{frame}

\begin{frame}
\frametitle{Assignment and Project Groups}
(I am just putting official first names here)
\begin{itemize}
\item Group 1: Fatemeh, Nazlinur, Agustinus
\item Group 2: Sondoss, Yasin, Taichi
\item Group 3: William, Haeyun, Timothy
\item Group 4: Afnan, Kristin
\item Group 5: Brody, Emily
\end{itemize}
\footnotesize Note: Contrary to what some of you thought, I actually did think about whom to put together. I wanted to put people with varying skill levels together (from my perspective), except for Group 5 (as they are the only two undergrads in the class).
\end{frame}


\begin{frame}
\frametitle{Assignment 5 description}
\begin{itemize}
\item Deadline: March 24th
\item Questions: 3 questions, Answer any 2. (I would strongly recommend not skipping Question 1)
\item Submission: One submission per group! 
\item Question 1: Using terminal in MacOS/Linux/Windows 10+ 
\item Question 2: Implementing a sorting function yourself
\item Question 3: Implementing a searching function yourself (This is more difficult than the other two)
\item You can do all 3 and get bonus credit of 2\% in the final grade, if you want. 
\end{itemize}
\end{frame}

\begin{frame}%CHANGE
\frametitle{Final Project: how is the grade split?}
\begin{itemize}
\item 30\% of your total grade
\item 10\% for a short oral presentation in the finals week.
\item 20\% for the code and a report, due on May 1st. (5\% for user interface, 10\% for other part of the code, 5\% for the report). 
\item Some project ideas on Canvas
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Discussion Time}
\begin{itemize}
\item Time for your ideas! (discuss with your team)
\item I will ask all teams to present their ideas in a few min. 
\item You should also post your idea (just a small paragraph) on the discussion forum (one post per team)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Week after break}
\begin{itemize}
\item Working with databases
\item databases + bottle
\item Assignment 5 will be due at the end of that week.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{}
\Large Happy Spring break!!
\end{frame}

\end{document}
