\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\author[Sowmya Vajjala]{Instructor: Sowmya Vajjala}

\title[ENGL 516]{ENGL 516: Methods of Formal Linguistic Analysis}
\subtitle{Semester: Spring '18}

\date{12 Apr 2018}

\institute{Iowa State University, USA}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frame}\titlepage
\end{frame}

\begin{frame}
\frametitle{Last class' exercise}
I want a program, that takes as input a .txt file path from the user, and prints the following: 
\begin{itemize}
\item Number of unique words/total number of words
\item Percentage of words in this document that appeared only once.
\item Average word frequency for top 25 words.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{}
\Large Solutions on the forum - discussion
\end{frame}

\begin{frame}
\frametitle{Project presentations schedule}
\begin{itemize}
\item 24th: 
\begin{enumerate}
\item Brody, Emily
\item Kristin, Afnan
\item Tim, Taichi, Fatemeh
\end{enumerate}
\item 26th:
\begin{enumerate}
\item Sondoss, Hardi, William
\item Yasin, Haeyun, Nazli
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Next class}
I compiled some readings which (I think) will give a perspective before starting 520. This time, I am going to pick three people to present these together with insights you may have obtained with 516 (time: 15 min max, + discussion): \small
\begin{itemize}
\item "The promise of NLP and speech processing technologies in language assessment" (LTJ, Chapelle and Chung, 2010)
\\ \footnotesize \url{http://journals.sagepub.com/doi/abs/10.1177/0265532210364405}
\\ (My chosen person: Fatemeh) \small
\item "Natural Language and Language Learning" (EAL, Meurers, 2013)
\\ \footnotesize \url{https://onlinelibrary.wiley.com/doi/pdf/10.1002/9781405198431.wbeal0858}
\\(My chosen person: Yasin) \small
\item Intelligent Computer Assisted Language Learning (EAL, Heift, 2013)
\\ \footnotesize \url{https://onlinelibrary.wiley.com/doi/pdf/10.1002/9781405198431.wbeal0548}
\\(My chosen person: William) \small
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Rest of today}
\begin{itemize}
\item You know know how to collect some basic corpus statistics.
\item Incorporate this into a web UI.
\item i.e., user uploads a text, and your program calculates these statistics (ttr, avg. word freq. for top 25 words, precentage of rare words) and shows back the output.
\item You have all the pieces in what we discussed earlier - you have to think through the approach, and how to put what you learnt so far together.
%\item If you think that is a waste of time - you can walk out, and don't show up in my class again. I will give you an A or A-.
\end{itemize}
\end{frame}

\end{document}
