\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{url}

\author[Sowmya Vajjala]{Instructor: Sowmya Vajjala}


\title[ENGL 516X]{ENGL 516X: \\ Methods of Formal Linguistic Analysis}
\subtitle{Semester: Spring '18}

\date{20 March 2018}

\institute{Iowa State University, USA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frame}\titlepage
\end{frame}

\begin{frame}
\frametitle{Announcements, Reminders etc.}
\begin{itemize}
\item Assignment 5 is due this weekend. 
\item Start working on your final project.
\item Try to spend more time on python in these last few weeks.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Questions about Last week}
\begin{itemize}
\item How many of you looked at the Web interface examples from last week? \pause
\item How many of you think you can develop a quick, simple interface in Python now? \pause
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Assignment 4 discussion}
\begin{itemize}
\item Answers seem to be slightly different among people (primarily because of different pre-processing)
\item Different formats of displaying outputs
\end{itemize}
-Like a nice teacher, I tolerated all these. So, I want 2 nice student volunteers - one for each question, to show their solutions.
\end{frame}

\begin{frame}
\frametitle{Topic for the week: Databases}
\begin{itemize}
\item What is a database? How can we work with it?
\item How can we "query" and modify databases with Structured Query Language?
\item How can we integrate a database and a UI in Python?
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{What is a database?}
\begin{itemize}
\item A database is a file (or files) used to save data in a structured manner. 
\item We can visualize it as a table (or collection of tables), with many rows and columns.
\item Each row is similar to a dictionary in python. There is a key, and a set of values associated with it. 
\item However, while a dictionary is something that is stored in a program and goes away when the program closes, databases are things that are stored on your hard drive.
\item Databases are a way to store and retrieve data quickly, and we can also make permanent changes to them in python programs.
\item Some popular examples: spreadsheets, comma separated value (csv) files etc.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Terminology}
\begin{itemize}
\item Database Management System: Software that allows users/programmers to interact with a database and work with the data in it.
\item Table, Rows, Columns
\item Keys: The unique identifier of a row in a table.
\item querying: issuing commands to a database to access or update data.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Basic operations with a database}
\begin{itemize}
\item Create: create a table or a collection of tables
\item inserting new rows or new columns into tables
\item updating (editing) an existing row in a table
\item deleting something from a table
\item selecting some rows from a table to view/print them or do something with them.
\item Developing relations between two or more tables by creating either common keys, or mapping one table's key to another etc.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Databases in python}
\begin{itemize}
\item we will use sqllite3 - a python library that will allow us to access a database stored on our hard disk and manipulate it.
\item To talk to the database, sqllite3 uses a format called "Structured Query Language (SQL)"  
\item To view a database, it is good to have an excel like software. In Firefox, a plugin called SQLLiteManager allows you to do that (or install SQLite browser from http://sqlitebrowser.org/ )
\item So, we write code to read/write info from database, and we can use SQLLiteManager to see if our code is really working.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{SQL Tutorial}
\begin{itemize}
\item Work in groups of 2-3 people, and do some lessons at: \url{http://www.w3schools.com/sql/}
\item In the left panel, you see lesson names. Do: SQL Intro, Syntax, Select, Where, Insert into, Update, Delete 
\item Let us say you spend about 30-40 min on this. 
\end{itemize}
\end{frame}

\begin{frame}%40min.
\frametitle{Additional practice material}
\begin{itemize}
\item \url{http://sqlzoo.net/}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Working with databases in Python}
\begin{itemize}
\item CreatingADBExample.py
\item InsertingIntoADBExample.py
\item PrintEditDeleteExamples.py
\end{itemize}
(I will continue with these examples on Thursday)
\end{frame}

\begin{frame}
\frametitle{Next Class}
\begin{itemize}
\item SQL in Python
\item TODO: Read Chapter 15 in textbook (There is a  lot of stuff in it, that I will not discuss, though)
\item TODO: Take a look at these code examples.
\end{itemize}
\end{frame}

\end{document}
