\documentclass{beamer}
\usepackage[utf8]{inputenc}

\author[Sowmya Vajjala]{Instructor: Sowmya Vajjala}


\title[ENGL 516X]{ENGL 516X: \\ Methods of Formal Linguistic Analysis}
\subtitle{Semester: Spring '18}

\date{22 Feb 2018}

\institute{Iowa State University, USA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frame}\titlepage
\end{frame}

\begin{frame}%2minutes
\frametitle{Class outline}
\begin{itemize}
\item Files: Wrap up
\item Dictionaries and Tuples in Python %10min
\item Tuples continued %30min
\item How to work with so many different data structures %10min
\item Other datastructures (general overview) %15min
\item Project ideas for final exam %15min
\end{itemize}
\end{frame}

\begin{frame}[fragile] %20min
\frametitle{Files: Last Class' Question}
\begin{itemize}
\item Reviewing ListFilesInDir.py \pause
\item Changing something there: using os.path.join() method.
\item Useful ref: \url{https://docs.python.org/3.1/library/os.path.html} \pause
\item Question: Counting number of lines in each file in a directory - solution: CountLinesOfFilesInDirectory.py on Canvas \pause
\item Did anyone try to use a pdf/doc file in open(filename).read()?
\end{itemize}
\end{frame} 

\begin{frame}[fragile] %20min
\frametitle{On the discussion forum}
\begin{itemize}
\item I am not discussing some of the solutions as there are multiple answers posted on the forum
\item Thanks to all the active participants there!
\item Please take a look once in a while, and try to understand other people's programs. Ask questions if you want. 
\item Reading and understanding other people's code is going to be very important once you stop doing classwork!
\end{itemize}
\end{frame} 

\begin{frame}%10min
\frametitle{Dictionaries in Python}
\begin{enumerate}
\item Dictionaries are a way of storing data as pairs (called: "key-value" pair). \pause
\item Think of a telephone directory - we access it by names (keys) to get numbers (values). 
\item An Example dictionary structure: eg = \{'Sowmya': 'India', 'Taichi': 'Japan', 'Nazlinur': 'Turkey', 'Brody':'USA'\} \pause
\\ In this, eg['Sowmya'] returns me 'India'.
\item A dictionary can have a dictionary embedded within itself too.
\item One good and bad thing about dictionaries: You don't access them one by one sequentially. i.e., you cannot do eg[1] etc
\item How does python know if we want a dictionary object? (Example code: xyz = dict())
\item Use: one example is to build a word frequency list from a corpus. 
\end{enumerate}
%introduce hashing
\end{frame}

\begin{frame}[fragile] %20min
\frametitle{How to work with dictionaries}
DictionaryBasics.py \tiny
\\ \url{https://docs.python.org/3/tutorial/datastructures.html#dictionaries}
\end{frame} 

\begin{frame}
\frametitle{Let us do a small exercise}
Exercise 3 in the Dictionaries Chapter
\\ "Write a program to read through a mail log, build a histogram using a dictionary to count how many messages have come from each email address, and print the dictionary"
\end{frame} 

\begin{frame}[fragile]
\frametitle{Dictionaries and Lists: questions}
\begin{itemize}
\item What is the value of someThing after this statement?
\begin{verbatim}
 someThing = {'a': 1,'b':2, 'a':3}
\end{verbatim} \pause
\item What does this print?
\begin{verbatim}
a=[1,2,3,4,5,6,7,8,9]
a[None:6]
\end{verbatim} \pause
\item What gets printed after the following two statements:
\begin{verbatim}
dict  = {"user","sowmya", "password","iastate"}
print(dict['sowmya'])
\end{verbatim}\pause
\item What does this print:
\begin{verbatim}
numbers = [1, 2, 3, 4]
numbers.append([5,6,7,8])
print(len(numbers))
\end{verbatim}
\end{itemize}
%Show a code and ask what it is doing and what the output is.
\end{frame}

\begin{frame}[fragile] %10min
\frametitle{Tuples: An Introduction}
\begin{itemize}
\item Tuples are a collection of items, that look very similar to lists in python.
\item However, the main difference is that tuples are immutable, whereas lists are mutable.
\item Dictionaries have a method called items(), that returns a list of "tuples", where each item is a key-value pair.
\begin{verbatim}
d = {'a':10, 'b':1, 'c':22}
t = d.items()
print(t)
[('a', 10), ('c', 22), ('b', 1)]
\end{verbatim}
\end{itemize} \tiny
\url{https://docs.python.org/3/tutorial/datastructures.html#tuples-and-sequences}
\end{frame} 

\begin{frame}[fragile]
\frametitle{Tuples and Dictionaries}
With tuples, we can have two iteration variables in a loop to read through a dictionary!
\begin{verbatim}
d = {'a':10, 'b':1, 'c':22}
for key, val in d.items():
    print(val, key)
#The output is:
10 a
22 c
1 b
\end{verbatim}
\end{frame}

\begin{frame}[fragile]
\frametitle{Defining a Tuple}
\begin{itemize}
\item t1 = 'a', 'b', 'c', 'd', 'e'
\item t2 = ('a', 'b', 'c', 'd', 'e')
\item t3 = tuple() \#creates empty tuple
\item t4 = ('a',)
\item lst = [1,2,3]
\item t5 = tuple(lst)
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Tuple Operations}
Several list and string operators work with tuples too. Let our example tuple be: t = ('a', 'b', 'c', 'd', 'e')
\begin{itemize}
\item t[0] gives us 'a'
\item slicing operation t[1:3] gives us a tuple ('b', 'c')
\item Something like: t[0] = 'A' will throw an error as tuples are immutable.
\item But, we can change the whole tuple. This won't throw an error: t = ('A',) + t[1:]
\item A very useful thing: (name,addr) = "sowmya@iastate.edu".split("@") - saves "sowmya" into name and "iastate.edu" in to addr.
\item name,addr = addr,name swaps name and addr values in just one line!
\end{itemize}
\end{frame}

\begin{frame}[fragile] %5min
\frametitle{Comparing Tuples}
\begin{itemize}
\item Comparison of tuples (and also other sequences like lists) works in a slightly counter-intuitive way. E.g., (0, 1, 2000000) $<$ (0, 3, 4) will be true in Python. \pause
\item Sort example:
\begin{verbatim}
txt = 'but soft what light in yonder window breaks'
words = txt.split()
t = list()
for word in words:
   t.append((len(word), word))
t.sort(reverse=True)
res = list()
for length, word in t:
    res.append(word)
print(res)
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile] %15min -2 slides
\frametitle{Back to Tuples and Dictionaries}
Sorting a dictionary by values.
\begin{verbatim}
d = {'a':10, 'b':1, 'c':22}
l = list()
for key, val in d.items():
    l.append((val, key))
print(l) #what will this print?
\end{verbatim} \pause
\begin{verbatim}
[(10, 'a'), (22, 'c'), (1, 'b')]
l.sort(reverse=True)
print(l)
\end{verbatim}
\end{frame}

\begin{frame}[fragile]
\frametitle{Tuples and Dictionaries -2}
Using tuples as dictionary keys: This is useful when we want to have dictionary keys as having more than one element. An example is: writing a program to create telephone directory.. where you want to index by both firstname and lastname. 
\begin{itemize}
\item Let us say we want to assign like this: directory[last,first] = number - (last,first) here is a tuple.
\item To write a for loop for this dictionary, we then write: \\ for last, first in dictionary: ...
\end{itemize}
\end{frame}

\begin{frame} %10 min
\frametitle{How to work with all these data structures}
\begin{itemize}
\item Lists, Dictionaries and Tuples can get very complex and confusing, especially if one is nested inside another. So, you should be careful while writing your programs and learn to effectively debug them. \pause
\item While debugging: when you are clueless about the errors you are getting, read the code line by line, work out manually for one or two examples to check if your program is doing everything right. \pause
\item Use try-except blocks to catch known issues. Have print statements at necessary places during the debugging phase, to know what the code is doing.\pause
\item Understand the different errors and why they occur. \pause
\item Keep experimenting and use online tutorials, search for advice on stackoverflow.com, look for video lectures.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Other Data Structures - 1} %15min total for two slides.
\begin{enumerate}
\item A Stack: Stack is a kind of data structure where we assume items as being stacked one above the other. So, whichever was put in last needs to be taken out first (this is called LIFO - Last in First Out). Example code and working: \url{http://goo.gl/g2W5pH} \pause
\item A Queue: Queue is the reverse of Stack. It is a FIFO (First in First Out). Example code and working: \url{http://goo.gl/Yw4zB0}
\end{enumerate}
\pause
Uses: internal implementation of search operations, performing arithmatic operations, recursive operations and many more. A good explanation of various ways to implement and apply stacks and queues is here: \url{https://goo.gl/2yA0OW}
\pause
\\ Both a stack and a queue can be implemented using the list data structure in Python. Try this out by writing your own code (call them Stack.py and Queue.py) for these. 
% point to: https://en.wikipedia.org/wiki/List_of_data_structures
\end{frame}

\begin{frame}
\frametitle{Other Data Structures - 2}
\begin{enumerate}
\item A Tree: A tree is a data structure with nodes and connections between them, arranged in a root, branch, leaves style.
\begin{itemize}
\item use: representing hierarchical information, storing data for searching through, for routing information from one location to another.
\item Trees also can be implemented using Lists in Python. Example: \url{http://goo.gl/UYE4yD}
\end{itemize} \pause
\item A Graph: This has a set of points called nodes or vertices and a lot of connections between them.
\begin{itemize}
\item use: any kind of problem which involves a network of relations (e.g., analysing social network, molecule interactions, computer networks 
finding the shortest route etc.,)
\item Graphs also can be implemented in Python using dictionaries and lists. Example: \url{http://goo.gl/3Tkjh2}
\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Exercise for today: post solutions later}
\begin{itemize}
\item Exercise 2 in textbook chapter: "Write a program that categorizes each mail message by which day of the week the commit was done. To do this look for lines that start with "From", then look for the third word and keep a running count of each of the days of the week. At the end of the program print out the contents of your dictionary (order does not matter)."
\item Note: You can use mbox-short.txt as the file to estimate these numbers. 
\item Hints: We need knowledge of- file processing, regular expressions, lists, dictionaries, strings to solve this. 
\end{itemize}
\end{frame}
%Practice exercises here.

\begin{frame}
\frametitle{Next Week}
\begin{itemize}
\item Tuesday: Review and Practice So, post any questions you have online in the forum with "Recap: 27 Feb" as the title.
\item Tuesday: I will prepare some exercises that will help you with Assignment 4
\item Thursday: Basics about making small applications that work on browsers.
\item Tasks for you: Revise and have questions. Post questions on forum by Monday.
\item Browse through this website: \url{https://bottlepy.org/docs/dev/}
\end{itemize}
\end{frame}

\begin{frame}%10min
\frametitle{Some Pythonic Fun}
"30 Python language features and tricks you may not know about" \\
\url{http://sahandsaba.com/thirty-python-language-features-and-tricks-you-may-not-know.html}
\end{frame}

\end{document}

