\documentclass[11pt,a4paper]{article}
\usepackage{hyperref}

\begin{document}
\begin{center}
  Spring Semester 2018 \\ Iowa State University\\[3ex]
  {\large ENGL 516 - Formal Methods of Linguistic Analysis}\\[3ex]
  \textbf{Assignment 6} \\ \textbf{Submission Deadline: 07 April 2018}
\end{center}


\paragraph{Instructions:} This assignment consists of one question, for 10 \% of your final grade. It involves creating a small web interface using Bottle, and creating and updating a database with user details. Save your code as groupnum\_assignment5.py. If the code does not run and throws errors, you cannot get a grade unless you defend it in the office hours. Weightage: 5\% for a working UI, 5\% for a working database logic.

\section*{Question 1}
Here is a description of what I want to see in your program: The purpose of this program is to register new users into a database of ALT enthusiasts. The opening page of your web interface should have a html page with the following fields:
a name field, that takes a name for the user. an email address field that takes an email address, an interests field - that takes a comma seperated string of the research interests of the user. When someone reaches your interface, you should prompt them to enter these details. Once they click on submit, you should go to your database, and: \\
a) If someone with that name already exists in your database, you should not do anything, but leave a message to the user that they need to use some other name, as someone with that name already exists. \\
b) If a new user (with a new name) arrives, you should add these details to your database, and leave a message to the user that their details have been added. 



\end{document} 
