\documentclass[11pt,a4paper]{article}
\usepackage{hyperref}

\begin{document}
\begin{center}
  Spring Semester 2018 \\ Iowa State University\\[3ex]
  {\large ENGL 516 - Formal Methods of Linguistic Analysis}\\[3ex]
  \textbf{Assignment 3} \\ \textbf{Submission Deadline: 17 Feb 2018}
\end{center}

\paragraph{Instructions:} This assignment consists of three questions. First 2 questions give you 2.5\% each, and 3rd question gives you 5\% of your final grade. All the questions involve writing python code. Save your code as three .py files (one for each question, in yourname\_q1.py, yourname\_q2.py etc format), and submit them as individual files on Canvas. If the program does not run and throws errors, you cannot get a grade unless you defend it in the office hours. Last question has a starter code, and you have to use it and expand it.

\begin{enumerate}
\item Write a program that takes a string as input, and loops through the string to print it in reverse. I don't want you to use any built in reversing functions in Python. I want you to write a loop.
\item Write a program that takes a number, and prints all numbers from 1 until that number, line by line. i.e., if I enter 5 as input, the program has to print:
1
2
3
4
\item This program takes two strings as input. One is a long string. The second one is a shorter one. It then prints the number of times the shorter string appears in the longer string. For example: let us say the longer string is: "baaabaa blaack sheep" and the short string is "aa", it should print 4 as output. Starter code is provided. Another example test case: if "Mississippi" and "ss" are inputs, it should return me 2 as output. 
\end{enumerate}

%string operations and loops.
%- number of overlapping substrings in a string (give two test cases, one overlapping; one non overlapping) (5
%- loop and print a string in reverse
%- write a program that takes a number and prints all numbers from 1 until that number (i.e., it should print 1, 2, 3, 4 line by line, if I enter 5 as input).

\end{document}
