\documentclass[10pt,a4paper]{article}
\usepackage{url}
\title{Practice exercises for the concepts from Chapter 4 and Chapter 5 in the textbook.}
\begin{document}
\maketitle

Source of the first two questions: \url{http://www.ling.gu.se/~lager/python_exercises.html}

\begin{enumerate}
\item A pangram is a sentence that contains all the letters of the English alphabet at least once, for example: The quick brown fox jumps over the lazy dog. Your task here is to write a function to check a sentence to see if it is a pangram or not.

\item "99 Bottles of Beer" is a traditional song in the United States and Canada. It is popular to sing on long trips, as it has a very repetitive format which is easy to memorize, and can take a long time to sing. The song's simple lyrics are as follows:
\\ 99 bottles of beer on the wall, 99 bottles of beer.
\\ Take one down, pass it around, 98 bottles of beer on the wall. \\
The same verse is repeated, each time with one fewer bottle. The song is completed when the singer or singers reach zero. Your task here is write a Python program capable of generating all the verses of the song.

\item Write a program with a function that asks the user how many Fibonnaci numbers to generate and then generates them. The Fibonnaci seqence is a sequence of numbers where the next number in the sequence is the sum of the previous two numbers in the sequence. The sequence looks like this: 1, 1, 2, 3, 5, 8, 13, and so on.  So, if the user enters 5, you should generate the first 5 Fibonacci numbers..i.e., 1,1,2,3,5
More on Fibonacci series: \url{https://www.mathsisfun.com/numbers/fibonacci-sequence.html}

\item Write a program with a function that checks whether a number is a prime number or not. More on prime numbers: \url{http://www.mathsisfun.com/definitions/prime-number.html}

\end{enumerate}
\end{document}
